---
title: ' HW2 '
Author: Hamza Kale
---
# 1. Introduction
In this homework, the ultimate task is to forecast the sales of UGS for every quarter of 2007 by using the data between 2000 and 2006.

## 1.1 Required Packages



```{r}
library(RcppRoll)
library(readxl)
library(lubridate)
library(zoo)
library(ggplot2)
library(scales)
library(data.table)
library(corrplot)
library(ggcorrplot)
library(GGally)
library(forecast)
library(dplyr)
```

### All Metrics:
UGS: Unleaded gasoline sale in a given quarter,
RNUV: An index indicating the rate of new unleaded gasoline using vehicles being added to the traffic in
a quarter,
PU: Average price (adjusted with an index) of a liter of unleaded gasoline in a quarter,
PG: Average price (adjusted with an index) of a liter of diesel gasoline in a quarter,
NUGV: Number of unleaded gasoline using vehicles in the traffic,
NDGV: Number of diesel gasoline using vehicles in the traffic (per 1000 people),
GNPA: Agriculture component of Gross National Product (adjusted with an index),
GNPC: Commerce component of Gross National Product (adjusted with an index),
GNP: Grand total for GNP (agriculture, commerce and other components total).

## 1.2 Data Acquisition
```{r}
path <- "C:/Users/asus/Desktop/IE360/HW2/IE360_Spring22_HW2_data.csv"

Data_HW2 <- read.csv(path,colClasses=c('character',rep('numeric',10)))
colnames(Data_HW2) <- c("Quarters", "UGS", "RNUV","NLPG","PU","PG","NUGV","NDGV","GNPA","GNPC","GNPT")

Data_HW2$Quarters <- as.Date(as.yearqtr(Data_HW2$Quarters, format = "%Y_Q%q"))


Data_Given <- Data_HW2[c(1:28),]
Data_Forecast <- Data_HW2[c(29,30,31,32),]

Data_Given <- data.table(Data_Given)
Data_Forecast <- data.table(Data_Forecast)

```



## 1.3 Visualization of the Data 2000-2006 and Results
```{r}
ggplot(Data_Given, aes(x=Quarters,y=UGS, group = 1)) +
geom_point() +
geom_line() +
labs(y ='Quarters')+ 
ggtitle('Unleaded Gasoline Sales (UGS) 2000-2006')
```
Results:
1) There is  a yearly seasonality in the data. 
2) There is a strongly decreasing trend.
3) The data is not stationary with respect to mean because of declining trend
4) The data is  stationary with respect to variance. With respect to this results, no transformation is unnecessary.


## 1.4 Autocorrelation Function of UGS
```{r}
acf(Data_Given$UGS)

```

In the autocorrelation plot, the values for lag 1 and lag 4 are above the limits. Because the lag 4 can be explained by quarterly seasonality, the only lagged variable which is added in the model is lag 1.

## 1.5 Correlation Diagram
```{r}
ggpairs(Data_Given)

```

According to correlation diagram above, attributes highly correlated with UGS are: 
1) NLPG
2) PU
3) PG
4) NUGV
5) NDGV
6) GNPA

# 2. FORECASTING

## 2.1 Adding Trend and Seasonality Components to the Model

```{r}
Data_Given[,trend := 1:.N ]
Data_Given[,Quarters_:=as.character(month(Quarters))]

Data_Forecast[,trend := 29:32 ]
Data_Forecast[,Quarters_:=as.character(month(Quarters))]
```


In the part of 1.3 Visualization of the Data 2000-2006 and Results, we conclude that there is  a yearly seasonality in the data and strongly decreasing trend. To optimize the model, seasonality and trend component is added.

```{r}

Model_2.1 <- lm(UGS ~ trend+Quarters_, Data_Given)
summary(Model_2.1)
```

Since the value of residual standard error is low and the value of R-squared is high, we prove that the new model is much better that the old one just by adding seasonality and trend components.

## 2.2 Adding the Correlated Variables
In the part of 1.5 Correlation Diagram, it is mentioned that some variables are correlated with UGS. Because of this correlations, the correlated variables should be added in the model.
```{r}
Model_2.2 <- lm(UGS ~ trend+NLPG+PU+PG+NUGV+NDGV+GNPA + Quarters_, Data_Given)
summary(Model_2.2)
```

After adding correlated variables the value of residual standard error becomes lower than part 2.1 and the value of R-squared increase by adding correlated variables. So it is proven that the correlated variable adding is an essential update to get better forecasts. However, the NUGV and NDGV variables are not important variables according to p-values of them.

## 2.3 Adding the Lagged Variable to the Model
```{r}
Data_Given$UGSlag1=lag(Data_Given$UGS,1)
Model_2.3 <-  lm(UGS ~ trend+PU+PG+NUGV+NDGV+GNPA +UGSlag1 + Quarters_, Data_Given)
summary(Model_2.3)

```

After adding lagged variable the value of residual standard error becomes lower than part 2.1 and the value of R-squared increase by adding lagged variable. So it is proven that the lagged variable adding is an essential update to get better forecasts.

## 2.4 Model Check
```{r}
checkresiduals(Model_2.3$residuals)

```

```{r}
Plot_Model2.3=copy(Data_Given)
Plot_Model2.3[,actual:=UGS]
Plot_Model2.3[,predicted_trend:=predict(Model_2.3,Plot_Model2.3)]
Plot_Model2.3[,residual_trend:=actual-predicted_trend]

ggplot(Plot_Model2.3 ,aes(x=Quarters)) +
        geom_line(aes(y=UGS,color='current data')) + 
        geom_line(aes(y=predicted_trend,color='forecast'))
```

As it seen above, the forecast approximately fits the real values. Therefore, Model2.3 can be used to forecast UGS values of 2007 quarters.

# 3. Forecasting 2007 UGS Values

1st Quarter - 655894.0
2nd Quarter - 850454.8
3rd Quarter - 961714.9
4th Quarter - 783118.0

```{r}

Data_Forecast$UGSlag1[1]=Data_Given$UGS[28]


Data_Forecast[1,"UGS"]=as.numeric(predict(Model_2.3,newdata=Data_Forecast[1,]))
Data_Forecast[2,"UGS"]=predict(Model_2.3,newdata=Data_Forecast[2,])
Data_Forecast[3,"UGS"]=predict(Model_2.3,newdata=Data_Forecast[3,])
Data_Forecast[4,"UGS"]=predict(Model_2.3,newdata=Data_Forecast[4,])

Data_Forecast$UGSlag1[2]=as.numeric(Data_Forecast[1,"UGS"])
Data_Forecast$UGSlag1[3]=as.numeric(Data_Forecast[2,"UGS"])
Data_Forecast$UGSlag1[4]=as.numeric(Data_Forecast[3,"UGS"])

Data_Forecast[,"UGS"]

```
1st Quarter - 655894.0
2nd Quarter - 850454.8
3rd Quarter - 961714.9
4th Quarter - 783118.0